---
title: "Crash course in R for FIE453"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

# Create student_df
n <- 100
student_df <- tibble(student_number = seq(1, n),
                    age = runif(n, 18, 40),
                    score = rnorm(n, 3, 1),
                    city = sample(c("Bergen", "Oslo", "Trondheim"), n, replace = T))

# Import german credit data
# Load data
url <- "http://freakonometrics.free.fr/german_credit.csv"
raw <- read_csv(url)

# Make names valid
credit_df <- raw %>%
  set_tidy_names(syntactic = TRUE)

# Select a subset of colums and rename for consistency
credit_df <- credit_df %>%
  select(Creditworthy = Creditability, 
         Duration_months = Duration.of.Credit..month., 
         Credit_amount = Credit.Amount, 
         Age = Age..years., 
         Occupation, 
         Foreign_worker = Foreign.Worker, 
         Gender_marital = Sex...Marital.Status) 

# Fix factor-columns
credit_df <- credit_df %>%
  mutate(
    Occupation = case_when(
      Occupation == 1 ~ "Unemployed - non-resident",
      Occupation == 2 ~ "Unemployed - resident",
      Occupation == 3 ~ "Normal worker",
      Occupation == 4 ~ "Manager or higher"
    ),
    Gender_marital = case_when(
      Gender_marital == 1 ~ "Man, single",
      Gender_marital == 2 ~ "Woman, married",
      Gender_marital == 3 ~ "Man, married",
      Gender_marital == 4 ~ "Woman, single"
    )
  )

```


## Topic 1 - basic R operations

Welcome to this crash course in R, which is created using the "learnr"-package in R. 

Please note that all exercises in this course are independent. That means that objects you create in one exercise will *not* be available in the next. 

### Vectors

*Let's start with the basics...*

Write the R code required to add two plus two:

```{r two-plus-two, exercise=TRUE}

```

Write the R code required to create a vector [2, 2] and name it "x". Print "x" to the console using print().

```{r vector, exercise=TRUE}

```

```{r vector-hint}
# Use the c-function to create vectors, e.g
x <- c(1, 1)
```

Write the R code required to create vectors y and z, y = [1, 2, 3] and z = [1, 3, 4]. 

1. Add y and z together.
2. *Combine* y and z, creating one long vector with 6 elements.
3. Create a new vector which tests if each element in y is equal to the corresponding element in z.
4. Compute the mean of y using the function mean().
5. Convert y to a character-vector. 

```{r vectorplus, exercise=TRUE}

```

```{r vectorplus-hint}
# The c-function can also be used to combine vectors
combined <- c(vector_a, vector_b)

# In order to convert the type of x to a character, use:
x <- as.character(x)

```

Create a new vector named "x_na", with the values [1, 3, NA]. "NA" here means that the value is missing.

1. Compute the mean of x_na. What do you get, and why?
2. Now compute the mean again, but add the option na.rm = TRUE to the function call. 


```{r missing, exercise=TRUE}

```

```{r missing-solution}

x_na <- c(1, 3, NA)

# The mean of an object with missing values defaults to "missing" in R.
mean(x_na) 

# In order to ignore the missing value, and compute the mean only based on existing numbers, use na.rm.
mean(x_na, na.rm = TRUE)

```

### Data frames (tibbles)

Create a tibble (note: a tibble is just an improved data frame) called "test_df" with columns "student_number" = [1, 2, 3], "age" = [23, 25, 28], "score" = [2, 3, 4], "city" = c("Bergen", "Oslo", "Trondheim"). Print the output using print()

Next, extract only the "age" column from test_df using "$" and print it. 

```{r dataframe, exercise=TRUE}
test_df <- tibble(student_number = ,
                  age = ,
                  score = ,
                  city = )
```


```{r dataframe-hint}
# To extract y from a datamframe named df, use $ as such:
df$y
```

Using the data frame format, it is easy to conduct analysis. We have already made a more complex version of *test_df* called *student_df*.

Run use a summary on the data frame called student_df:

```{r dataframe_summary, exercise=TRUE}

```

In order to illustrate the usefulness of putting data in a data frame, let's look at the relationship between age and grade. For now, just run the code below - we will come back to plotting later on. 

```{r dataframe_plot, exercise=TRUE}
ggplot(student_df, aes(x = age, y = score)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~ city) +
  theme_minimal()
```

### Quiz
```{r quiz}
quiz(
  question("What is the difference between a list and a data frame?",
    answer("A data frame can contain anything, while a list is limited to equal length vectors"),
    answer("A list can contain anything, while a data frame is limited to equal length vectors", correct = TRUE),
    answer("A data frame is simply a more efficient type of list"),
    answer("A list is simply a more efficient type of data frame")
  ),
  question("Can a vector contain two elements of different data types?",
    answer("No", correct = TRUE),
    answer("Yes"),
    answer("It depends")
  ),
  question("What is the difference between == and = in R?",
    answer(" = can be used for assignment, while == is used for logical comparisons", correct = TRUE),
    answer("They are equal, but == is more commonly used for logical comparisons"),
    answer(" = is not used at all in R, while == is used for logical comparisons")
  ),
    question("What does it mean that R is a vectorized language?",
    answer("R is optimized for working with vector operations, as opposed to e.g. loops", correct = TRUE),
    answer("R can only use vectors, and everything that exists in R is either a vector or based on a vector")
  )
)
```

## Topic 2 - data manipulation
Now we will move on to use real data, more specifically a data set from a German credit bank. The data has been imported already, named "credit_df".

Run a summary on the data frame. 

```{r summary2, exercise = TRUE}

```

Select only the "Foreign_worker"-column from "credit_df". 

```{r select, exercise = TRUE}

```

```{r select-hint}
# If you have a dataframe, df, you can select column y as follows:
df %>%
  select(y)

```

Create a new variable which is TRUE if Foreign_worker == 1 and FALSE otherwise. Save it to credit_df, then use the table() - function on the new variable to study its distribution. 

```{r mutate, exercise = TRUE}

```

```{r mutate-solution}
# In order to add a variable to a dataframe, you can use mutate as such:
credit_df <- credit_df %>%
  mutate(is_foreign = if_else(Foreign_worker == 1, TRUE, FALSE))

# if_else first computes a TRUE/FALSE statement, and assigns the first value after the comma to all 
# observations  where the expression evaluates to TRUE. 

table(credit_df$is_foreign)

```

Create a new data frame, "credit_df_filtered", which only contains observations where Creditworthy is 1. How many rows are left in this new dataframe, and how many percent of the observations are creditworthy?

```{r filter, exercise = TRUE}

```

```{r filter-hint}
# In order to filter a dataframe, use the filter-function.
df_filtered <- df %>%
  filter(x > 3)

# Count rows in each df
nrow(df_filtered) / nrow(df)
```

Find the average age for customers where Creditworthy is 1 and 0 respectively. Use group_by + summarise. 

```{r summarise, exercise = TRUE}

```

```{r summarise-hint}
# We can do grouped calculations in summarise after using group by.
# For example, lets calculate the median credit amount for those who are and are not creditworthy. 
credit_df %>%
  group_by(Creditworthy) %>%
  summarise(median_credit_amount = median(Credit_amount))

```

We have a variable named Gender_marital which is a combination of gender and marital status. Use table() to find out how many observations we have of each instance. 

```{r table, exercise = TRUE}

```

Create a new variable, Gender, which is either "male" or "female", using if_else(). Find out how many customers who are male. 

```{r ifelse, exercise = TRUE}

```

```{r ifelse-solution}
# You could create the variable as such. Note that the %in%-operator tests if a vector belongs to another vector
credit_df <- credit_df %>%
  mutate(Gender = if_else(Gender_marital %in% c("Man, married", "Man, single"), "male", "female"))

credit_df %>%
  count(Gender)
```

Now, create a variable which divides Credit_amount in three groups: < 1000, 1000 - 2000 and > 2000. You can use the function *case_when*, which is similar to if_else but for can be used in cases with more than 2 options.

```{r cut, exercise = TRUE}

```

```{r cut-hint}
# In order to use case_when to split up a variable, say age, in 3 groups, do as such:
df <- df %>%
  mutate(age_group = case_when(age < 18 ~ "Child",
                               age < 70 ~ "Adult",
                               TRUE ~ "Senior"))

# The expression reads as follows: first, everyone under 18 gets tagged as a "child", 
# then everyone who are under 70 gets tagged as adult. Finally, the only ones who are left will be seniors,
# who are tagged in the last statement, which tags everyone who isn't tagged yet with the label "senior". 

```

Lastly, find the average age and credit amount for all different values of occupation, but only for foreign workers.

```{r bonus, exercise = TRUE}

```

```{r bonus-hint}
# You need a filter, a group_by and a summarise! 

```

```{r quiz_data_manipulation}
quiz(
  question(
    "What does the function mutate do?",
    answer("It splits a variable into two columns."),
    answer("It adds one or more variables to a tibble", correct = TRUE),
    answer("It is used to join together two different tibbles")
  ),
  question(
    "Which package is the most used for data manipulation in R?",
    answer("dplyr", correct = TRUE),
    answer("plyr"),
    answer("zoo"),
    answer("caret")
  ),
  question(
    "What is a left_join?",
    answer("left_join(x, y) keeps all obs. in y and adds columns based on x"),
    answer("left_join(x, y) keeps all obs. in x and adds columns based on y", correct = T)
  ),
  question(
    "What does the following code do: df %>% filter(x > 3)",
    answer("It removes all observations where x > 3"),
    answer("It keeps all observations where x > 3", correct = T)
  )
)
```

## Topic 3 - data visualization with ggplot
### Distribution plots
Plot a histogram for the "Age"-variable in credit_df. Choose your favorite color!

```{r histogram, exercise = TRUE}

```

```{r histogram-hint}
# In order to plot a histogram of the variable bar from a dataframe df you write:
ggplot(df, aes(x = bar)) +
  geom_histogram(color = "black", fill = "pink")

# Choose color and fill as you like.

```

In the same histogram, set bins = 20.
```{r histogrambins, exercise = TRUE}

```

Now replace "geom_histogram" with "geom_density".

```{r density, exercise = TRUE}

```

In the same density plot, choose a different *fill* for the different groups of Creditworthy. Consider using the alpha-parameter to make the plots easier to see. NB! You might have to convert Creditworthy to a factor first.

```{r densityfill, exercise = TRUE}

```

```{r densityfill-solution}
ggplot(credit_df, aes(x = Age, fill = factor(Creditworthy))) +
  geom_density(alpha = 0.5)
```

### Relationship between variables
Use a box-plot (geom_boxplot) to illustrate which occupations that have the highest credit amount. Bonus: add a title to the plot.

```{r boxplot, exercise = TRUE}

```

```{r boxplot-hint}
# In order to create a boxplot of y (continious variable) w.r.t x (factor):
ggplot(df, aes(x = x, y = y)) +
  geom_boxplot()

# For adding titles, see ?labs

```

Illustrate the relationship between Creditworthy and Occupation. Choose a suitable *geom* (this one is a bit harder...)

```{r barplot, exercise = TRUE}

```

```{r barplot-hint}
# You could start by finding the average creditiability per occupation (group_by + summarise)
# This could be plotted using e.g. a bar plot, geom_bar. 
```

Next, illustrate the relationship between length of credit and occupation. 

```{r scatter, exercise = TRUE}

```

```{r scatter-hint}
# Since we here are looking at the relationship between two numerical variables, you could e.g. use a scatter plot combined with a trend, as such:
ggplot(df, aes(x = x, y = y)) +
  geom_point() +
  geom_smooth()

```

Create the same plot, except by coloring the different groups of Creditworthy in different colors. 

```{r scattercolor, exercise = TRUE}

```

Create the same plot, but now add a third variable: split the plot by occupation. Use facet_wrap. Feel free to experiment making your plot as beautiful as possible - pick suitable colors, pick a theme, add a header etc. 

```{r facet, exercise = TRUE}

```

```{r facet-hint}
# Plot x against y, split by z
ggplot(df, aes(x = x, y = y)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~ z)
```

```{r quiz_viz}
quiz(
  question("When is it useful to use geom_point (e.g. a scatter plot)",
    answer("When you are plotting a numeric variable against a factor variable"),
    answer("When you are plotting two numeric variables against each other", correct = TRUE),
    answer("When you are plotting two factor variables")
  ),
  question("When is it useful to use geom_boxplot?",
    answer("When you are plotting a numeric variable against a factor variable", correct = TRUE),
    answer("When you are plotting two numeric variables against each other"),
    answer("When you are plotting two factor variables")
  ),
  question("What is the difference between geoms and aesthetics?",
    answer("geoms indicate which variables should be plotted, aesthetics indicate the visuals to be used"),
    answer("aesthetics indicate which variables should be plotted, geoms indicate the visuals to be used", correct = TRUE)
  )
)
```
